Integrating Common Ground and Informativeness in Pragmatic Word Learning

### 2019-04-15 papers/cogsci2019/cogsci2019.Rmd
clean up repo

```{r}
# data ex 1 and ex 2
  d1 <- bind_rows(read_csv(file="../../stats/data/novel.data.csv"),
               read_csv(file="../../stats/data/pref.data.csv")) %>%
  mutate(trial_type = ifelse(trial == "train", "train", "test"),
         control = ifelse(change == TRUE, "true" , "false"))

d2 <- read_csv(file="../../stats/data/inf.data.csv") %>%
  mutate(trial_type = ifelse(trial == "train", "train", "test"),
         control = ifelse(control == TRUE, "true", "false"))

dex12 <- bind_rows(d1,d2) %>%
  filter(trial_type == "test")%>%
  mutate(control= relevel(factor(control), ref = "true"),
         choiceAgent = ifelse( control == "false", agent, altAgent)) # re-code agent for preference and novelty

# data ex 3
d3 <- bind_rows(read_csv(file="../../stats/data/ex3.novel.data.csv"),
               read_csv(file="../../stats/data/ex3.pref.data.csv")) %>%
  mutate(trial_type = ifelse(trial == "train1" | trial =="train2", "train", "test"))

# check if someone did both experiments
# yes, two people
x <- d3 %>%
  filter(trial_type == "train")%>%
  group_by(id)%>%
  summarise(n = length(correct_inf)) %>%
  filter(n > 2)

# exclude those who did both experiments
dex3 <- d3 %>%
  filter(!id %in% x$id) %>%
  mutate(Speaker = ifelse(change =="same", "Same speaker", "Different speaker"),
         Alignment = ifelse(alignment == "congruent", "Congruent", "Incongruent"),
         Prior = ifelse(experiment == "pref_inf", "Preference", "Novelty"),
         choiceAgent = ifelse (change == "same", agent, altAgent))%>%
  filter(trial_type == "test")

d4 <- bind_rows(read_csv(file="../../stats/data/ex3.2.novel.strong.data.csv"),
               read_csv(file="../../stats/data/ex3.2.novel.medium.data.csv"),
               read_csv(file="../../stats/data/ex3.2.novel.weak.data.csv"),
               read_csv(file="../../stats/data/ex3.2.pref.strong.data.csv"),
               read_csv(file="../../stats/data/ex3.2.pref.medium.data.csv")
               ) %>%
  mutate(trial_type = ifelse(trial == "train1" | trial =="train2", "train", "test"))

x <- d4 %>%
  filter(trial_type == "train") %>%
  group_by(id)%>%
  summarise(correct_inf = mean(correct_inf)) %>%
  filter(correct_inf == 0)

dex4 <- d4 %>%
  filter(!id %in% x$id)%>% 
  distinct(id, alltrial, .keep_all = TRUE) %>%
  filter(trial_type != "train") %>%
  mutate(Change = ifelse(change =="same", "Same speaker", "Different speaker"),
         Alignment = ifelse(alignment == "congruent", "Congruent", "Incongruent"),
         Experiment = ifelse(grepl("pref",experiment), "Preference", "Novelty"),
         Prior = ifelse(prior == "weak", "Weak", ifelse(prior == "medium", "Medium", "Strong"))
         )

@@

```{r ex1-results}
ex1t <- dex12 %>%
  filter(condition == "informativeness")%>%
  group_by(control, id) %>%
  summarise(correct = mean(correct)) %>%
  summarise(correct = list(correct)) %>%
  group_by(control) %>%
  mutate(df= t.test(unlist(correct), mu = 0.5)$parameter,
         t_value = t.test(unlist(correct), mu = 0.5)$statistic,
         p_value = t.test(unlist(correct), mu = 0.5)$p.value) %>%
  select(control,df,t_value,p_value)%>%
  mutate(p_value = ifelse(p_value<.001,"< .001",as.character(paste("=",substr(round(p_value,3),2,5),sep=" "))),
         t_value = round(t_value,2))

# model
lm_ex1 <- glmer(correct ~ control +
              (control |id) + (control |agent), 
              data = dex12%>%filter(condition == "informativeness"), 
              family = binomial,control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

@@

Participants selected the less frequent object above chance in the test condition (t(`r ex1t%>%filter(control == "false")%>%pull(df)`) = `r ex1t%>%filter(control == "false")%>%pull(t_value)`, *p* `r ex1t%>%filter(control == "false")%>%pull(p_value)`, see Fig. \ref{fig:plotexp12}) and did so more often compared to the control condition (generalized linear mixed model (GLMM\footnote{All models included random effects for subject and speaker.}):*$\beta$* = `r ex1_r%>%filter(term == "controlfalse")%>%pull(estimate)`, se = `r ex1_r%>%filter(term == "controlfalse")%>%pull(std.error)` *p* `r ex1_r%>%filter(term == "controlfalse")%>%pull(p.value)`). This result replicates earlier work [@frank2014inferring] and is consistent with the hypothesis that listeners expect speakers to communicate in an informative way.

@@

p1 <- dex12 %>%
  filter(trial_type == "test") %>%
  mutate(control = ifelse(control == "true", "Control", "Test"),
         condition = factor(condition), 
         experiment = ifelse(condition == "informativeness", "1","2"))%>%
  group_by(experiment,condition ,control, id) %>%
  summarise(correct = mean(correct))

@@

 ggplot() +
  geom_jitter(data = p1, aes(x = control, y = correct, col = control), alpha = 0.2, width = .3,height = .025)+
  geom_pointrange(data = p2, aes(x = control, y = mean, col = control,ymin = ci_lower, ymax = ci_upper),size = .5)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Expected Choice")+
  facet_wrap( ~ condition, labeller = as_labeller(c(`preference`="Exp. 2 - Pref.", `novelty`="Exp. 2 - Nov.", `informativeness`="Exp. 1 - Inf.")))+
  theme_few(base_size = 10) + 
  ylim(-0.05,1.05)+
  guides(alpha = F,size =F, col = F)+
  scale_color_manual(name = "Condition",
                     breaks=c("Test","Control"),
                     values= c("#839496","black"))

@@

```{r ex2-results-pref}
ex2tpref <- dex12 %>%
  filter(condition == "preference")%>%
  group_by(control, id) %>%
  summarise(correct = mean(correct)) %>%
  summarise(correct = list(correct)) %>%
  group_by(control) %>%
  mutate(df= t.test(unlist(correct), mu = 0.5)$parameter,
         t_value = t.test(unlist(correct), mu = 0.5)$statistic,
         p_value = t.test(unlist(correct), mu = 0.5)$p.value) %>%
  select(control,df,t_value,p_value)%>%
  mutate(p_value = ifelse(p_value<.001,"< .001",as.character(paste("=",substr(round(p_value,3),2,5),sep=" "))),
         t_value = round(t_value,2))

# model (maximally converging)
lm_ex2pref <- glmer(correct ~ control +
              (1|id) + (control |choiceAgent), 
              data = dex12%>%filter(condition == "preference"), 
              family = binomial,control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

@@

```{r ex2-results-nov}
ex2tnov <- dex12 %>%
  filter(condition == "novelty")%>%
  group_by(control, id) %>%
  summarise(correct = mean(correct)) %>%
  summarise(correct = list(correct)) %>%
  group_by(control) %>%
  mutate(df= t.test(unlist(correct), mu = 0.5)$parameter,
         t_value = t.test(unlist(correct), mu = 0.5)$statistic,
         p_value = t.test(unlist(correct), mu = 0.5)$p.value) %>%
  select(control,df,t_value,p_value)%>%
  mutate(p_value = ifelse(p_value<.001,"< .001",as.character(paste("=",substr(round(p_value,3),2,5),sep=" "))),
         t_value = round(t_value,2))

# model
lm_ex2nov <- glmer(correct ~ control +
              (1|id) + (control |choiceAgent), 
              data = dex12%>%filter(condition == "novelty"), 
              family = binomial,control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

@@

Participants selected the preferred object above chance when the same animal returned (t(`r ex2tpref%>%filter(control == "false")%>%pull(df)`) = `r ex2tpref%>%filter(control == "false")%>%pull(t_value)`, *p* `r ex2tpref%>%filter(control == "false")%>%pull(p_value)`, see Fig. \ref{fig:plotexp12}) and did so more often compared to trials in which a different animal returned (GLMM: $\beta$ = `r ex2pref_r%>%filter(term == "controlfalse")%>%pull(estimate)`, se = `r ex2pref_r%>%filter(term == "controlfalse")%>%pull(std.error)`, *p* `r ex2pref_r%>%filter(term == "controlfalse")%>%pull(p.value)`). Thus, listeners inferred the referent of the utterance by considering previous interactions with the speaker. Interestingly, participants transferred preference to some extent from one animal to the other and selected the preferred object above chance when a different animal returned (t(`r ex2tpref%>%filter(control == "true")%>%pull(df)`) = `r ex2tpref%>%filter(control == "true")%>%pull(t_value)`, *p* `r ex2tpref%>%filter(control == "true")%>%pull(p_value)`). In sum, this study shows that adults make comparable inferences to children [@saylor2009preschoolers].

The novel object was selected above chance when the same animal returned (t(`r ex2tnov%>%filter(control == "false")%>%pull(df)`) = `r ex2tnov%>%filter(control == "false")%>%pull(t_value)`, *p* `r ex2tnov%>%filter(control == "false")%>%pull(p_value)`) but not when a different one appeared (t(`r ex2tnov%>%filter(control == "true")%>%pull(df)`) = `r ex2tnov%>%filter(control == "true")%>%pull(t_value)`, *p* `r ex2tnov%>%filter(control == "true")%>%pull(p_value)`, see Fig. \ref{fig:plotexp12}). Furthermore, the two conditions differed in the expected direction ($\beta$ = `r ex2nov_r%>%filter(term == "controlfalse")%>%pull(estimate)`, se = `r ex2nov_r%>%filter(term == "controlfalse")%>%pull(std.error)`, *p* `r ex2nov_r%>%filter(term == "controlfalse")%>%pull(p.value)`). Thus, like children [@akhtar1996role], adults used their prior information about the speaker to resolve ambiguity in the utterance.

@@

p3 <- dex3 %>%
  group_by(Prior,Speaker ,Alignment, id) %>%
  summarise(correct = mean(correct_inf))

@@

ggplot() +
  #geom_rect(data = p4, aes(fill = NA), color = "#859900", size = 2.5, alpha = 0.4 ,xmin = -Inf,xmax = Inf, ymin = -Inf,ymax = Inf)+
  geom_jitter(data = p3, aes(x = Alignment, y = correct), alpha = 0.2, width = .3,height = .025)+
  geom_pointrange(data = p4, aes(x = Alignment, y = mean,ymin = ci_lower, ymax = ci_upper),size = .5)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Choosing More Informative")+
  facet_grid ( ~ Prior + Speaker) +
  theme_few(base_size = 10) + 
  ylim(-0.05,1.05)+
  guides(alpha = F, fill = F, col = F)+ 
  scale_fill_manual(name="Condition",
                     breaks=c("1","2","3"),
                     values= c("#859900","#859900","#859900"))

@@

```{r ex3_tab, results="asis"}
#comaprison to chance
ex3_tab <- dex3 %>%  
  filter(trial_type == "test") %>%
  group_by(Speaker ,Prior,Alignment, id) %>%
  summarise(correct_inf = mean(correct_inf)) %>%
  summarize(correct_inf = list(correct_inf)) %>%
  group_by(Speaker ,Prior,Alignment) %>%
  mutate(M = mean(unlist(correct_inf)),
         W = wilcox.exact(unlist(correct_inf), mu = 0.5)$statistic,
         p = wilcox.exact(unlist(correct_inf), mu = 0.5)$p.value) %>%
  select(Prior,Alignment, Speaker, M , W,p)%>%
  mutate(M = round(M,2),
         p = ifelse(p<.001,"< .001",as.character(substr(round(p,3),2,5),sep=" ")))

@@

```{r ex3_results, cache=T}
# model 
lm_ex3 <- glmer(correct_inf ~ Prior*Speaker*Alignment + (Speaker+Alignment|id) + (Speaker+Alignment|choiceAgent), 
              data = dex3, family = binomial, 
              control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

@@

Results are discussed in the form of the proportion with which listeners chose the more informative object (i.e., the object that would be the more informative referent when only considering speaker general expectations). For a comparison to chance within each unique condition see Fig. \ref{fig:plotexp3}. Combinations of alignment and speaker identity differed in how they influenced participants’ responses (GLMM model term: `alignment*speaker`; $\beta$ = `r lm_ex3_r%>%filter(term == "SpeakerSame speaker:AlignmentIncongruent")%>%pull(estimate)`, se = `r lm_ex3_r%>%filter(term == "SpeakerSame speaker:AlignmentIncongruent")%>%pull(std.error)`, *p* `r lm_ex3_r%>%filter(term == "SpeakerSame speaker:AlignmentIncongruent")%>%pull(p.value)`). Fig. \ref{fig:plotmodelcomp}A shows the mean response in each unique condition compared to the pragmatics model. Model predictions and data were highly correlated (r = `r model_comp_r$estimate`, *p* `r model_comp_r$p.value`). Model fit was much better in the model taking into account both types of expectations compared to the uniform prior (BF = `r format(bf_ex3%>%filter(Comparison == "Pragmatic vs. No prior") %>%pull(BayesFactor),digits = 2)`) or prior only model (BF = `r format(bf_ex3%>%filter(Comparison == "Pragmatic vs. Prior only") %>%pull(BayesFactor), digits = 2)`). The inferred noise level in the pragmatics model was 0.27 (95% HDI: 0.21 - 0.34).

@@

In our model, we treated common ground expectations as equivalent to more basic manipulations of contextual salience [e.g. in @frank2012predicting] and did not explicitly model the social-cognitive processes that give rise to these expectations. The interaction around the object prior to the test event simply increased the probability that this particular speaker will refer to the object subsequently. The same change could be brought about if one of the objects would be made perceptually more salient, for example by making it flash. In future work, it would be interesting to explore ways to model common ground expectations explicitly as well as to contrast perceptual and interactional salience.

@@

This work integrates different perspectives on the study of pragmatic inference. Previous work focused either on general or speaker specific expectations. The methodological approach taken here illustrates how computational and experimental approaches can be used in conjunction to explicate theories of language use and learning.

@@

\vspace{1em} \fbox{\parbox[b][][c]{7.3cm}{\centering Corresponding data and code are available at\ \url{https://github.com/manuelbohn/mcc}}} \vspace{1em} -->

@@

Masked for peer review.

>>>>

```{r}
# data ex 1 and ex 2

adult_ex1_data <- read_csv(file="../../stats/data/adult_ex1.csv")
# adults exp 2 novelty
adult_ex2_novelty_data <- read_csv(file="../../stats/data/adult_ex2_novelty.csv")
# adults exp 2 preference
adult_ex2_preference_data <- read_csv(file="../../stats/data/adult_ex2_preference.csv")

dex12 <- bind_rows(
    adult_ex1_data,
    adult_ex2_novelty_data %>% mutate(condition = ifelse(condition == "same_speaker", "test", "control")),
    adult_ex2_preference_data %>% mutate(condition = ifelse(condition == "same_speaker", "test", "control")))

# data ex 3
dex3 <- read_csv(file="../../stats/data/adult_ex3.csv")

# data ex 4

dex4 <- read_csv(file="../../stats/data/adult_ex4.csv")

@@

```{r ex1-results}
ex1t <- dex12 %>%
  filter(experiment == "adult_ex1")%>%
  group_by(condition, id) %>%
  summarise(correct = mean(correct)) %>%
  summarise(correct = list(correct)) %>%
  group_by(condition) %>%
  mutate(df= t.test(unlist(correct), mu = 0.5)$parameter,
         t_value = t.test(unlist(correct), mu = 0.5)$statistic,
         p_value = t.test(unlist(correct), mu = 0.5)$p.value) %>%
  select(condition,df,t_value,p_value)%>%
  mutate(p_value = ifelse(p_value<.001,"< .001",as.character(paste("=",substr(round(p_value,3),2,5),sep=" "))),
         t_value = round(t_value,2))

# model
lm_ex1 <- glmer(correct ~ condition +
              (1 |id), 
              data = dex12%>%filter(experiment == "adult_ex1"), 
              family = binomial,control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

@@

Participants selected the less frequent object above chance in the test condition (t(`r ex1t%>%filter(condition == "test")%>%pull(df)`) = `r ex1t%>%filter(condition == "test")%>%pull(t_value)`, *p* `r ex1t%>%filter(condition == "test")%>%pull(p_value)`, see Fig. \ref{fig:plotexp12}) and did so more often compared to the control condition (generalized linear mixed model (GLMM\footnote{All models had maximal random effects structure conditional on model convergence.}):*$\beta$* = `r ex1_r%>%filter(term == "conditiontest")%>%pull(estimate)`, se = `r ex1_r%>%filter(term == "conditiontest")%>%pull(std.error)` *p* `r ex1_r%>%filter(term == "conditiontest")%>%pull(p.value)`). This result replicates earlier work [@frank2014inferring] and is consistent with the hypothesis that listeners expect speakers to communicate in an informative way.

@@

p1 <- dex12 %>%
  mutate(condition = ifelse(condition == "control", "Control", "Test"),
         condition = factor(condition))%>%
  group_by(experiment,condition , id) %>%
  summarise(correct = mean(correct))

@@

 ggplot() +
  geom_jitter(data = p1, aes(x = condition, y = correct, col = condition), alpha = 0.2, width = .3,height = .025)+
  geom_pointrange(data = p2, aes(x = condition, y = mean, col = condition,ymin = ci_lower, ymax = ci_upper),size = .5)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Expected Choice")+
  facet_wrap( ~ experiment, labeller = as_labeller(c(`adult_ex2_preference`="Exp. 2 - Pref.", `adult_ex2_novelty`="Exp. 2 - Nov.", `adult_ex1`="Exp. 1 - Inf.")))+
  theme_few(base_size = 10) + 
  ylim(-0.05,1.05)+
  guides(alpha = F,size =F, col = F)+
  scale_color_manual(name = "Condition",
                     breaks=c("Test","Control"),
                     values= c("#839496","black"))

@@

```{r ex2-results-pref}
ex2tpref <- dex12 %>%
  filter(experiment == "adult_ex2_preference")%>%
  group_by(condition, id) %>%
  summarise(correct = mean(correct)) %>%
  summarise(correct = list(correct)) %>%
  group_by(condition) %>%
  mutate(df= t.test(unlist(correct), mu = 0.5)$parameter,
         t_value = t.test(unlist(correct), mu = 0.5)$statistic,
         p_value = t.test(unlist(correct), mu = 0.5)$p.value) %>%
  select(condition,df,t_value,p_value)%>%
  mutate(p_value = ifelse(p_value<.001,"< .001",as.character(paste("=",substr(round(p_value,3),2,5),sep=" "))),
         t_value = round(t_value,2))

# model (maximally converging)
lm_ex2pref <- glmer(correct ~ condition +
              (1|id), 
              data = dex12%>%filter(experiment == "adult_ex2_preference"), 
              family = binomial,control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

@@

```{r ex2-results-nov}
ex2tnov <- dex12 %>%
  filter(experiment == "adult_ex2_novelty")%>%
  group_by(condition, id) %>%
  summarise(correct = mean(correct)) %>%
  summarise(correct = list(correct)) %>%
  group_by(condition) %>%
  mutate(df= t.test(unlist(correct), mu = 0.5)$parameter,
         t_value = t.test(unlist(correct), mu = 0.5)$statistic,
         p_value = t.test(unlist(correct), mu = 0.5)$p.value) %>%
  select(condition,df,t_value,p_value)%>%
  mutate(p_value = ifelse(p_value<.001,"< .001",as.character(paste("=",substr(round(p_value,3),2,5),sep=" "))),
         t_value = round(t_value,2))

# model
lm_ex2nov <- glmer(correct ~ condition +
              (condition|id), 
              data = dex12%>%filter(experiment == "adult_ex2_novelty"), 
              family = binomial,control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

@@

Participants selected the preferred object above chance when the same animal returned (t(`r ex2tpref%>%filter(condition == "test")%>%pull(df)`) = `r ex2tpref%>%filter(condition == "test")%>%pull(t_value)`, *p* `r ex2tpref%>%filter(condition == "test")%>%pull(p_value)`, see Fig. \ref{fig:plotexp12}) and did so more often compared to trials in which a different animal returned (GLMM: $\beta$ = `r ex2pref_r%>%filter(term == "conditiontest")%>%pull(estimate)`, se = `r ex2pref_r%>%filter(term == "conditiontest")%>%pull(std.error)`, *p* `r ex2pref_r%>%filter(term == "conditiontest")%>%pull(p.value)`). Thus, listeners inferred the referent of the utterance by considering previous interactions with the speaker. Interestingly, participants transferred preference to some extent from one animal to the other and selected the preferred object above chance when a different animal returned (t(`r ex2tpref%>%filter(condition == "control")%>%pull(df)`) = `r ex2tpref%>%filter(condition == "control")%>%pull(t_value)`, *p* `r ex2tpref%>%filter(condition == "control")%>%pull(p_value)`). In sum, this study shows that adults make comparable inferences to children [@saylor2009preschoolers].

The novel object was selected above chance when the same animal returned (t(`r ex2tnov%>%filter(condition == "test")%>%pull(df)`) = `r ex2tnov%>%filter(condition == "test")%>%pull(t_value)`, *p* `r ex2tnov%>%filter(condition == "test")%>%pull(p_value)`) but not when a different one appeared (t(`r ex2tnov%>%filter(condition == "control")%>%pull(df)`) = `r ex2tnov%>%filter(condition == "control")%>%pull(t_value)`, *p* `r ex2tnov%>%filter(condition == "control")%>%pull(p_value)`, see Fig. \ref{fig:plotexp12}). Furthermore, the two conditions differed in the expected direction ($\beta$ = `r ex2nov_r%>%filter(term == "conditiontest")%>%pull(estimate)`, se = `r ex2nov_r%>%filter(term == "conditiontest")%>%pull(std.error)`, *p* `r ex2nov_r%>%filter(term == "conditiontest")%>%pull(p.value)`). Thus, like children [@akhtar1996role], adults used their prior information about the speaker to resolve ambiguity in the utterance.

@@

p3 <- dex3 %>%
  mutate(common_ground_manipulation = paste0(toupper(substr(common_ground_manipulation, 1, 1)), substr(common_ground_manipulation, 2, nchar(common_ground_manipulation))),
         speaker = paste0(toupper(substr(speaker, 1, 1)), substr(speaker, 2, nchar(speaker))),
         speaker = str_replace(speaker, "_", " "),
         alignment = paste0(toupper(substr(alignment, 1, 1)), substr(alignment, 2, nchar(alignment))))%>%
  group_by(common_ground_manipulation,speaker ,alignment, id) %>%
  summarise(correct = mean(correct_inf))

@@

ggplot() +
  #geom_rect(data = p4, aes(fill = NA), color = "#859900", size = 2.5, alpha = 0.4 ,xmin = -Inf,xmax = Inf, ymin = -Inf,ymax = Inf)+
  geom_jitter(data = p3, aes(x = alignment, y = correct), alpha = 0.2, width = .3,height = .025)+
  geom_pointrange(data = p4, aes(x = alignment, y = mean,ymin = ci_lower, ymax = ci_upper),size = .5)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Choosing More Informative")+
  facet_grid ( ~ common_ground_manipulation + speaker) +
  theme_few(base_size = 10) + 
  ylim(-0.05,1.05)+
  guides(alpha = F, fill = F, col = F)+ 
  scale_fill_manual(name="Condition",
                     breaks=c("1","2","3"),
                     values= c("#859900","#859900","#859900"))

@@

```{r ex3_tab, results="asis"}
#comaprison to chance
ex3_tab <- dex3 %>%  
  group_by(speaker ,common_ground_manipulation,alignment, id) %>%
  summarise(correct_inf = mean(correct_inf)) %>%
  summarize(correct_inf = list(correct_inf)) %>%
  group_by(speaker ,common_ground_manipulation,alignment) %>%
  mutate(M = mean(unlist(correct_inf)),
         W = wilcox.exact(unlist(correct_inf), mu = 0.5)$statistic,
         p = wilcox.exact(unlist(correct_inf), mu = 0.5)$p.value) %>%
  select(speaker ,common_ground_manipulation,alignment, M , W,p)%>%
  mutate(M = round(M,2),
         p = ifelse(p<.001,"< .001",as.character(substr(round(p,3),2,5),sep=" ")))

@@

```{r ex3_results, cache=F}
# model 
lm_ex3 <- glmer(correct_inf ~ common_ground_manipulation*speaker*alignment + (alignment|id), 
              data = dex3, family = binomial, 
              control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

@@

Results are discussed in the form of the proportion with which listeners chose the more informative object (i.e., the object that would be the more informative referent when only considering speaker general expectations). For a comparison to chance within each unique condition see Fig. \ref{fig:plotexp3}. Combinations of alignment and speaker identity differed in how they influenced participants’ responses (GLMM model term: `alignment*speaker`; $\beta$ = `r lm_ex3_r%>%filter(term == "speakersame_speaker:alignmentincongruent")%>%pull(estimate)`, se = `r lm_ex3_r%>%filter(term == "speakersame_speaker:alignmentincongruent")%>%pull(std.error)`, *p* `r lm_ex3_r%>%filter(term == "speakersame_speaker:alignmentincongruent")%>%pull(p.value)`). Fig. \ref{fig:plotmodelcomp}A shows the mean response in each unique condition compared to the pragmatics model. Model predictions and data were highly correlated (r = `r model_comp_r$estimate`, *p* `r model_comp_r$p.value`). Model fit was much better in the model taking into account both types of expectations compared to the uniform prior (BF = `r format(bf_ex3%>%filter(Comparison == "Pragmatic vs. No prior") %>%pull(BayesFactor),digits = 2)`) or prior only model (BF = `r format(bf_ex3%>%filter(Comparison == "Pragmatic vs. Prior only") %>%pull(BayesFactor), digits = 2)`). The inferred noise level in the pragmatics model was 0.27 (95% HDI: 0.21 - 0.34).

@@

A range probabilistic frameworks been used to model word learning [e.g. @fazly2010probabilistic; @frank2009using; @xu2007word]. RSA models differ from other approaches in that they treat word learning as the outcome of a social reasoning process. In contrast to models for cross-situational word learning [@fazly2010probabilistic; @frank2009using], RSA models show how learning might occur in a one shot scenario based on pragmatic reasoning alone. While the ad hoc informativeness inference characteristic for RSA would be predicted by the model of Xu and Tenenbaum [-@xu2007word], in their work it follows from the "size principle" of generalization [@tenenbaum2001generalization] and not from social reasoning. In cotrast to RSA, this approach does not offer a straightforward way to incorporate other types of social information such as expectations following from common ground.

We treated common ground expectations as equivalent to more basic manipulations of contextual salience [e.g. in @frank2012predicting] and did not explicitly model the social-cognitive processes that give rise to these expectations. The interaction around the object prior to the test event simply increased the probability that this particular speaker will refer to the object subsequently. The same change could be brought about if one of the objects would be made perceptually more salient, for example by making it flash. In future work, it would be interesting to explore ways to model common ground expectations explicitly as well as to contrast perceptual and interactional salience.

@@

Our work integrates different perspectives on the study of pragmatic inference. Previous work focused either on general or speaker specific expectations. The methodological approach taken here illustrates how computational and experimental approaches can be used in conjunction to explicate theories of language use and learning.

@@

\vspace{1em} \fbox{\parbox[b][][c]{7.3cm}{\centering Corresponding data and code are available at\ \url{https://github.com/manuelbohn/mcc}}} \vspace{1em}

@@
